---
output:
  reprex::reprex_document:
    venue: "gh"
    advertise: FALSE
    session_info: TRUE
    style: TRUE
    comment: "#;-)"
    tidyverse_quiet: FALSE
    std_out_err: TRUE
knit: reprex::reprex_render
---

# Unexpected Behavior with `terra::split` Function

**Issue Overview:**

I am encountering unexpected behavior while attempting to split a polygon into two separate polygons using the `terra::split` function. In the following example, I am using the Bilag.shp dataset from [this repository](https://github.com/derek-corcoran-barrios/SplitError) and trying to split a specific arm of a fjord using a line. The split operation does not behave as intended and extends further than expected.

**Reprex:**

```{r}
# Load necessary libraries
library(terra)
library(ggplot2)
library(tidyterra)

# Load the dataset and create the splitting line
PartialBilag <- terra::vect("Bilag.shp")

latitudes <- c((55 + 56.245/60), (55 + 56.719/60))
longitudes <- c((11 + 54.364/60), (11 + 54.279/60))

Border_Roskilde <- data.frame(lon = longitudes, lat = latitudes) |>
  terra::vect(geom = c("lon", "lat"), crs = "+proj=longlat +datum=WGS84") |>
  terra::project("EPSG:25832") |>
  terra::as.lines()

# Visualize the data
ggplot() +
  geom_spatvector(data = PartialBilag, fill = "blue") +
  geom_spatvector(data = Border_Roskilde, color = "red") +
  theme_bw()
```

## Problematic Behavior:

The issue arises when I attempt to use the terra::split function to split the polygon using the defined line:

```{r}
Splited_Bilag <- terra::split(PartialBilag, Border_Roskilde)
Splited_Bilag$Polygon <- c("A", "B")

ggplot() + geom_spatvector(data = Splited_Bilag, aes(fill = Polygon)) + theme_bw()
```

*Observation:*

As you can see, the split operation extends the line southwards, covering a larger area than intended. The expected behavior was for the split to occur precisely along the defined line, rather than extending further south.

*Desired Functionality:*

In my understanding, the terra::split function should only utilize the specified line for the splitting operation, without extending the split southwards. I would appreciate clarification on whether this is the intended functionality or if there's a workaround to achieve the desired outcome.

Thank you for your attention and assistance in resolving this matter.

Best regards,

Derek
